---
title: "Global STP Analysis - S. pistillata"
author: "Matt"
date: "November 16, 2014"
output: html_document
---

### Load required libraries

```{r setup}
library("phyloseq")
library("ggplot2")
library("plyr")
library("vegan")
library("grid")
library("directlabels")
library("knitr")
setwd("./data")
opts_knit$set(root.dir = "./data", fig.keep='high')
opts_chunk$set(dev = 'pdf', fig.path="figures/sea/")
```

### Extract out seawater samples

```{r}
microSubRelFiltSea <- subset_samples(microSubRelFilt, species=='seawater')
```

# Relative abundance of seawater otus bar plot

```{r, fig.width = 10}
microSubRelFiltSeaFilt = filter_taxa(microSubRelFiltSea, function(x) mean(x) > 5e-3, TRUE)
seaBar <- plot_bar(microSubRelFiltSeaFilt, fill="Genus")
seaBar + facet_wrap(~site, scales="free")
```

Mostly dominated by Synechococccus and other typical seawater bacteria. Note the difference for Ningaloo - these were the samples that got warm! Seems this caused the community to change toward Alteromonas / Vibrio. 

# Check which Endozoicomonas OTUs are in the seawater

```{r}
microSubRelFiltSeaEndo = subset_taxa(microSubRelFiltSea, Genus=='Endozoicomonas(100)')
seaBarEndo <- plot_bar(microSubRelFiltSeaEndo, fill="Genus", title='Stylophora pistillata')
seaBarEndo + facet_wrap(~site, scales='free')

# add coloring for different Endozoicomonas OTUs
tax_table(microSubRelFiltSeaEndo) <- cbind(tax_table(microSubRelFiltSeaEndo), Strain=taxa_names(microSubRelFiltSeaEndo))
myranks = c("Genus", "Strain")
mylabels = apply(tax_table(microSubRelFiltSeaEndo)[, myranks], 1, paste, sep="", collapse="_")
tax_table(microSubRelFiltSeaEndo) <- cbind(tax_table(microSubRelFiltSeaEndo), catglab=mylabels)

microSubRelFiltSeaEndoFilt = filter_taxa(microSubRelFiltSeaEndo, function(x) mean(x) > 1e-10, TRUE)

plot_bar(microSubRelFiltSeaEndo, fill="catglab", title='seawater') +
  facet_wrap(~site, scales='free')




microSubRelFiltSpistEndoFilt = filter_taxa(microSubRelFiltSpistEndo, function(x) mean(x) > 1e-4, TRUE)
plot_bar(microSubRelFiltSpistEndoFilt, fill="catglab", title='Stylophora pistillata') +
  facet_wrap(~site, scales='free')
```





