---
title: "Global STP Analysis - Stylophora pistillata"
author: "Matt"
date: "November 16, 2014"
output: html_document
---

### Load required libraries


```{r setup, message=FALSE}
library("phyloseq")
library("ggplot2")
library("plyr")
library("vegan")
library("grid")
library("directlabels")
library("knitr")
library("clustsig")
library("ellipse")
setwd("./data")
opts_knit$set(root.dir = "./data", fig.keep='high')
opts_chunk$set(dev = 'pdf', fig.path="figures/spist/")
```

### Import data into R studio

First .shared 'otu matrix' from mothur

```{r}
sharedFile = read.table("micro.final.shared")
sharedFile = t(sharedFile)
rownames(sharedFile) = sharedFile[,1]
colnames(sharedFile) = sharedFile[2,]
sharedFile = sharedFile[,2:234]
sharedFile = sharedFile[4:37368,]
class(sharedFile) <- "numeric"
```

Import subsampled otu matrix (7779 seqs)

```{r}
sharedsubFile = read.table('micro.final.clean.0.03.subsample.shared')
sharedsubFile = t(sharedsubFile)
rownames(sharedsubFile) = sharedsubFile[,1]
colnames(sharedsubFile) = sharedsubFile[2,]
sharedsubFile = sharedsubFile[,2:219]
sharedsubFile = sharedsubFile[4:14839,]
class(sharedsubFile) <- "numeric"
```

Import taxonomy file from mothur

```{r}
taxFile = read.table('micro.final.0.03.cons.taxonomy', header=T, sep='\t')
rownames(taxFile) = taxFile[,1]
taxFile = taxFile[,2:8]
taxFile = as.matrix(taxFile)
```

Import metadata - ie. site, type, etc..

```{r}
metaFile = read.table('pools.metaData2', header=T, sep='\t')
rownames(metaFile) = metaFile[,1]
metaFile = metaFile[,2:7]
```

### Create phyloseq object

```{r}
OTU = otu_table(sharedFile, taxa_are_rows = TRUE)
OTUsub = otu_table(sharedsubFile, taxa_are_rows = TRUE)
TAX = tax_table(taxFile)
META = sample_data(metaFile)
physeq = phyloseq(OTU, TAX, META)
physeqSub = phyloseq(OTUsub, TAX, META)
physeqSub
```

Get rid of any OTUs not present in any samples and get relative abundance

```{r, fig.width = 4}
microSub <- prune_taxa(taxa_sums(physeqSub) > 0, physeqSub)
microSubRel = transform_sample_counts(microSub, function(x) x / sum(x) )
microSubRelFilt = filter_taxa(microSubRel, function(x) mean(x) > 1e-5, TRUE)
```

### Stylophora pistillata from aquaria

Take a look at the bacteria classified to genus and to family. NOTE: have to do this on non-subsampled data as these aqarium samples get removed in this step because of low sequence numbers (~3000)

```{r, fig.width = 4}
spistAq = prune_samples(c("SPaq", "SPaq-rep2"), physeq)
spistAqrelAbund = transform_sample_counts(spistAq, function(x) x / sum(x) )
spistAqrelAbundFilt = filter_taxa(spistAqrelAbund, function(x) mean(x) > 1e-2, TRUE)
plot_bar(spistAqrelAbundFilt, fill="Genus")
plot_bar(spistAqrelAbundFilt, fill="Family")
```

Pretty cool! Still about 1/4 of the bacteria are Endozoicomonas

Ok let's split the dataset into the S.pistillata samples, excluding replicates and the warmed ningaloo samples

```{r}
microSubRelFiltSpist <- prune_samples(c("11","49-2","50-2","51-2","52-2","53-2","54-2","57-2","63-2","64-2","65-2","67-2","80-2","75-2","76-2","77-2","101","102","103","105","173","174","175","176","205","206","207","208","209","226","227","228","229","230","231","235","236","237","241","242","243","244","251","253","254","256","257","258","259","260","261","262","MIC-25","MIC-50","MIC-53","MIC-91","MIC-145","MIC-151","MIC-220","MIC-226","MIC-229","MIC-424","MIC-427","MIC-453","RS1","RS2","RS3","RS5","RS6","RS7","RS10","RS12","RS13","RS15","RS16","RS17","RS18","19-rep2","55-2-rep2","56-2-rep2","104-rep2","MIC-85-rep2"), microSubRelFilt)
```

### Relative abundance of otus in all Stylophora pistillata samples

```{r, fig.width = 10}
microSubRelFiltSpistFilt = filter_taxa(microSubRelFiltSpist, function(x) mean(x) > 1e-2, TRUE)
spistBar <- plot_bar(microSubRelFiltSpistFilt, fill="Genus", title='Stylophora pistillata')
spistBar
```

Some variability between the samples but plenty of Endozoicomonas. I'll facet out the sites to take a closer look. 

```{r, fig.width= 10, fig.height = 6}
microSubRelFiltSpistFilt = filter_taxa(microSubRelFiltSpist, function(x) mean(x) > 1e-2, TRUE)
spistBar <- plot_bar(microSubRelFiltSpistFilt, fill="Genus", title='Stylophora pistillata')
spistBar + facet_wrap(~site, scales='free')
```

Quite some variability in the microbial composition across the sites. Can see Endozoicomonas throughout the graph, though it's a little difficult to see. I'll redo the graph with just endozoicomonas OTUs.

```{r, fig.width = 10, fig.height = 6}
microSubRelFiltSpistEndo = subset_taxa(microSubRelFiltSpist, Genus=='Endozoicomonas(100)')
spistBarEndo <- plot_bar(microSubRelFiltSpistEndo, fill="Genus", title='Stylophora pistillata')
spistBarEndo + facet_wrap(~site, scales='free')
```

Again quite a bit of variability between and within sites but can see Endozoicomonas at all the sites! Note the change in y-axis numbers (relative abundance) across the different sites - at the American Samoan sites this number is very small. 

The Red Sea has the most samples and is a bit hard to see at this scale - I'll take a closer look at these reefs.

```{r, fig.width= 10, fig.height = 8}
microSubRelFiltSpistFilt = filter_taxa(microSubRelFiltSpist, function(x) mean(x) > 1e-2, TRUE)
microSubRelFiltSpistFiltRedSea <- subset_samples(microSubRelFiltSpistFilt, site=="RedSea")
spistRedSea <- plot_bar(microSubRelFiltSpistFiltRedSea, aes(x=factor(Sample)), fill="Genus", title='Stylophora pistillata in the Red Sea')
spistRedSea + facet_wrap(~reef, scales="free_x")
spistRedSea + facet_wrap(~morphology, scales="free_x")
```

Endozoicomonas are fairly abundant throughout all sites, though this is variable. 

### Ordination of Stylophora pistillata samples

Which distance measure to use?

```{r, fig.width = 10, fig.height = 10, warning=FALSE}
dist_methods <- unlist(distance("list"))
dist_methods <- dist_methods[-(1:2)]
dist_methods <- dist_methods[-which(dist_methods == "ANY")]

plist <- vector("list", length(dist_methods))
names(plist) = dist_methods
for (j in dist_methods) {
  iDist <- distance(microSubRelFiltSpistFilt, method = j)
  iMDS <- ordinate(microSubRelFiltSpistFilt, "MDS", distance = iDist)
  p <- NULL
  p <- plot_ordination(microSubRelFiltSpistFilt, iMDS, color="site")
  p <- p + ggtitle(paste("MDS using distance method ", j, sep=""))
  plist[[j]] = p
}

df = ldply(plist, function(x) x$data)
names(df)[1] <- "distance"
p = ggplot(df, aes(Axis.1, Axis.2, color=site))
p = p + geom_point(size = 2, alpha = 0.5)
p = p + facet_wrap(~distance, scales = "free")
p = p + ggtitle("MDS on various distance metrics")
p

```

Pretty hard to decide on a distance metric I think - most people use Bray Curtis so guess I'll use that - even though I'm pretty sure Pat Schloss (the mothur guy) doesn't like this measure. 

Use the braycurtis matrix to generate plots using nMDS and PCoA

```{r, fig.width = 8}
mothurNMDS <- as.data.frame(read.table("micro.spist.0.03.braycurtis.0.03.lt.nmds.axes", header=T))
mothurPCOA <- as.data.frame(read.table("micro.spist.0.03.braycurtis.0.03.lt.pcoa.axes", header=T))

mothurNMDSspearman <- as.data.frame(read.table("micro.spist.0.03.spearman.corr.axes", header=T))
rownames(mothurNMDS) <- mothurNMDS[,1] 
rownames(mothurPCOA) <- mothurPCOA[,1]
rownames(mothurNMDSspearman) <- mothurNMDSspearman[,1]

mothurNMDSmeta <- merge(mothurNMDS, metaFile, by = "row.names")
rownames(mothurNMDSmeta) <- mothurNMDSmeta[,1]
mothurPCOAmeta <- merge(mothurPCOA, metaFile, by = "row.names")

ggplot(mothurPCOAmeta, title='PCoA of Stylophora pistillata samples') +
  geom_point(aes(x=axis2, y=axis3, color=site)) 

mothurNMDSplot <- ggplot(mothurNMDSmeta, title='nMDS of Stylophora pistillata samples') +
  geom_point(aes(x=axis2, y=axis3, color=site))
mothurNMDSplot

mothurNMDSplot + geom_polygon(aes(x=axis2, y=axis3, fill = site), alpha=0.5) + ggtitle("nMDS of Stylophora pistillata samples")
```

I think the nMDS looks better - it explains more of the variation and spreads the samples better.

Modify shapes so can display the different reefs within the sites

```{r, fig.width = 8}
mothurNMDSmeta.shape.names = unique(mothurNMDSmeta$reef)
mothurNMDSmeta.shape <- 1:(length(mothurNMDSmeta.shape.names))
names(mothurNMDSmeta.shape) <- mothurNMDSmeta.shape.names
mothurNMDSmeta.shape["samples"] <- 23
mothurNMDSplot <- ggplot(mothurNMDSmeta, title='nMDS of Stylophora pistillata with reef symbols') +
  geom_point(aes(x=axis2, y=axis3, color=site, shape = reef)) +
  scale_shape_manual(values = mothurNMDSmeta.shape) +
  theme(legend.key.size=unit(0.3,"cm"))
mothurNMDSplot
```

Actually I think the grouping is much stronger by site rather than by reef. A few of the Ningaloo reefs group together but the Red Sea reefs seem to be all mixed together. 

We can also overlay this plot with the OTUs that most contributed to the sample differences (generated using the indicator command in mothur) or by the spearman correlation

```{r, fig.width = 8}
top10indicator = mothurNMDSspearman[c("Otu000004","Otu000007","Otu000019","Otu000045","Otu000095","Otu000384","Otu000502","Otu000035","Otu000032","Otu000015","Otu000142"),]

top10spearman = mothurNMDSspearman[c("Otu000001","Otu000019","Otu000054","Otu000005","Otu000007","Otu000110","Otu000045","Otu000142","Otu000014","Otu000031","Otu000062", "Otu000018"),]

arrowmatrix = top10spearman
arrowdf <- data.frame(labels = rownames(arrowmatrix), arrowmatrix)

# get taxonomic information from the original tax file
taxFile[rownames(arrowmatrix),"Genus"]
arrowdf <- data.frame(labels = taxFile[rownames(arrowmatrix),"Genus"], arrowmatrix)

arrowmap <- aes(xend = axis2, yend = axis3, x = 0, y = 0, alpha=0.5, shape = NULL, color = NULL, label = labels)
labelmap <- aes(x = axis2, y = axis3 + 0.04, shape = NULL, color = NULL, label = labels, size=1.5)
arrowhead = arrow(length = unit(0.02, "npc"))
mothurNMDSplot <- ggplot(mothurNMDSmeta) +
  geom_point(aes(x=axis2, y=axis3, color=site))
mothurNMDSplotArrow <- mothurNMDSplot + geom_segment(arrowmap, size = 0.5, data = arrowdf, color = "black",  arrow = arrowhead, show_guide = FALSE) + geom_text(labelmap, size = 3, data = arrowdf)
mothurNMDSplotArrow 
```

This result reflects nicely what was in the barplots. Looks like the Ningaloo corals have a different Endozoicomonas OTU than the Red Sea reefs which is highly correlated with the samples. I'll look deeper into different Endozoicomonas types (oligotypes) across sites later. 

Also can do a SIMPER analysis on these matrices to check how much variation particular OTUs explain

```{r}
spistSharedSIMP = read.table('micro.spist.0.03.shared', header=TRUE)
rownames(spistSharedSIMP) <- spistSharedSIMP[,2]
spistSharedSIMP = spistSharedSIMP[,4:4813]
spistDistSIMP <- merge(spistSharedSIMP, metaFile, by="row.names")
rownames(spistDistSIMP) <- spistDistSIMP[,1]
OTUsSIMP <- spistDistSIMP[,2:4811]
factorsSIMP <- spistDistSIMP[,4812:4816]

spistSIMP <- simper(OTUsSIMP, factorsSIMP$site)
```

Lots of OTUs coming up here - plenty of Endozoicomonas OTUs are important in differentiating the sites. 

### Correlations between chemical data and microbes

Import nutrient and FCM data and look at Spearman correlations between this data and the sample ordinations

```{r, fig.width = 8}
Hbact <- as.data.frame(read.table("meta.Hbact.spearman.corr.axes", header=T))
NH4 <- as.data.frame(read.table("meta.NH4.spearman.corr.axes", header=T))
NN <- as.data.frame(read.table("meta.NN.spearman.corr.axes", header=T))
NO2 <- as.data.frame(read.table("meta.NO2.spearman.corr.axes", header=T))
PEPeuk <- as.data.frame(read.table("meta.PE+Peuk.spearman.corr.axes", header=T))
Peuk <- as.data.frame(read.table("meta.PEUK.spearman.corr.axes", header=T))
po4 <- as.data.frame(read.table("meta.po4.spearman.corr.axes", header=T))
PRO <- as.data.frame(read.table("meta.PRO.spearman.corr.axes", header=T))
silicate <- as.data.frame(read.table("meta.silicate.spearman.corr.axes", header=T))
SYN <- as.data.frame(read.table("meta.SYN.spearman.corr.axes", header=T))

nutrients <- rbind(NH4, NN, NO2, po4, silicate)
FCM <- rbind(Hbact, PEPeuk, Peuk, PRO, SYN)

arrowmatrix = nutrients
arrowdf <- data.frame(labels = arrowmatrix$Feature, arrowmatrix)

arrowmap <- aes(xend =axis2, yend = axis3, x = 0, y = 0, alpha=0.5, shape = NULL, color = NULL, label = labels)
labelmap <- aes(x = axis2, y = axis3 + 0.04, shape = NULL, color = NULL, label = labels, size=1.5)
arrowhead = arrow(length = unit(0.02, "npc"))
mothurNMDSplot <- ggplot(mothurNMDSmeta) +
  geom_point(aes(x=axis2, y=axis3, color=site))
mothurNMDSplotArrow = mothurNMDSplot + geom_segment(arrowmap, size = 0.5, data = arrowdf, color = "black",  arrow = arrowhead, show_guide = FALSE) + geom_text(labelmap, size = 3, data = arrowdf)
mothurNMDSplotArrow 

arrowmatrix = FCM
arrowdf <- data.frame(labels = arrowmatrix$Feature, arrowmatrix)

arrowmap <- aes(xend =axis2, yend = axis3, x = 0, y = 0, alpha=0.5, shape = NULL, color = NULL, label = labels)
labelmap <- aes(x = axis2, y = axis3 + 0.04, shape = NULL, color = NULL, label = labels, size=1.5)
arrowhead = arrow(length = unit(0.02, "npc"))
mothurNMDSplot <- ggplot(mothurNMDSmeta) +
  geom_point(aes(x=axis2, y=axis3, color=site))
mothurNMDSplotArrow = mothurNMDSplot + geom_segment(arrowmap, size = 0.5, data = arrowdf, color = "black",  arrow = arrowhead, show_guide = FALSE) + geom_text(labelmap, size = 3, data = arrowdf)
mothurNMDSplotArrow 
```

PRO = Prochlorococcus  
SYN = Synechococcus  
PEUK = Photosynthetic Eukaryotes  
PE+PEUK= Photosynthetic Eukaryotes that have a high PE fluorescence and are high in Chl as well  
HBact = non-pigmented bacteria  

Note in these plots that we don't have data for Indonesia or micronesia (maybe Amy has some), so the vectors cannot point toward those sites and may be a bit misleading - I could draw this with those sites excluded. Anyhow, the Red Sea sites have higher ammonia, while Ningaloo had more silicate and NO2. 

For the FCM results, the Red Sea had more Synechococcus and less Prochlorococcus. 

### SIMPROF analysis to check which samples fall into 'groups' without any *a priori* assumptions

Need to import the shared file containing just spist OTUs, then calcualte the simprof clusters based on the braycurtis metric. 

```{r, fig.width = 10, warning=FALSE}
spistShared = read.table('micro.spist.0.03.shared')
spistShared = t(spistShared)
rownames(spistShared) = spistShared[,1]
colnames(spistShared) = spistShared[2,]
spistShared = spistShared[,2:81]
spistShared = spistShared[4:4813,]
class(spistShared) <- "numeric"

spistSIMPROF <- simprof(spistShared, num.expected=10, num.simulated=9, method.cluster='average', method.distance='braycurtis', method.transform='identity', alpha=0.05, sample.orientation='column', silent=FALSE)

simprof.plot(spistSIMPROF, leafcolors=NA, plot=TRUE, fill=TRUE, leaflab="perpendicular", siglinetype=1)
```

Ok, pretty cool but bit difficult to tell which groups are falling out. I'll try and overlay the significant clusters on top of the nMDS. 

After calculating the clusters, make a data frame of the results and add to previous nMDS plot. Need to add these groups to the nMDS data.frame - I'll do a loop for this.

```{r, fig.width = 9, fig.height = 7, warning=FALSE}

simprofCLUSTERS = data.frame()

for(j in 1:length(spistSIMPROF$significantclusters)){
  if(length(spistSIMPROF$significantclusters[[j]]) > 2){
    simprofCLUSTERS <- rbind(simprofCLUSTERS, cbind(j, spistSIMPROF$significantclusters[[j]]))
  }
}

rownames(simprofCLUSTERS) <- simprofCLUSTERS[,2]
colnames(simprofCLUSTERS) <- c("simprofCLUSTER", "group")
mothurNMDSmetaSIMP <- merge(mothurNMDSmeta, simprofCLUSTERS, by="row.names")

#plot over nMDS

df_ell <- data.frame()
for(g in levels(mothurNMDSmetaSIMP$simprofCLUSTER)){
  df_ell <- rbind(df_ell, cbind(as.data.frame(with(mothurNMDSmetaSIMP[mothurNMDSmetaSIMP$simprofCLUSTER==g,], ellipse(cor(axis2, axis3), scale=c(sd(axis2), sd(axis3)), centre=c(mean(axis2), mean(axis3))))),site=g))
}

nMDSsimprof <- ggplot(data=mothurNMDSmeta, aes(x=axis2, y=axis3, color=site)) +
  geom_point(size=2, alpha=0.6) +
  scale_color_hue(limits=levels(droplevels(mothurNMDSmeta$site)))
nMDSsimprof

nMDS.mean <- aggregate(df_ell, list(df_ell$site), FUN=mean)

#write.table(df_ell, file="simpBrayIdent.txt")
#write.table(nMDS.mean, file="simpBrayIdentMean.txt")

df_ell_saved <- (read.table("simpBrayIdent.txt"))
df_ell_saved$site <- factor(df_ell_saved$site)
nMDS.mean.saved <- read.table("simpBrayIdentMean.txt")

nMDSsimprof +
  geom_path(data=df_ell_saved, aes(x=x, y=y, color=site), size=0.5, linetype=2, show_guide=FALSE) +
  annotate("text", x=nMDS.mean.saved$x, y=nMDS.mean.saved$y, label=nMDS.mean.saved$Group.1)
```

Looks a bit busy but actually some pretty cool clusters coming out here. I went back to the spreadsheet and photos to check out if there were any patterns to the clusters. 


mothurNMDSmetaSIMP[mothurNMDSmetaSIMP$simprofCLUSTER==6,"group.x"]

**Clusters**  
**3** - 103  RS12 RS15 RS16 RS17 RS18 **Rita's sites RW and RK**  
**5** - RS1 RS2 RS3 RS5 RS6 RS7 **Rita's sites RA and RJ**  

**7** - 52-2  MIC-85-rep2 RS13 **Ningaloo 1, Woleai, RK**  
**10** - 75-2 76-2 77-2 **all Ningaloo site 4, normal looking Stylophora**  
**11** - 11 49-2 50-2 51-2 53-2 57-2 63-2 64-2 65-2 **Ningaloo sites 1-3, normal looking Stylophora**  

**6** - 101     MIC-151 MIC-229 **Am Soa, Woleai, Olimareo**  
**13** - 19-rep2 226     230     259 **Al Lith, Inner Fsar, Maggie Is**  
**18** - 229     237     256     MIC-220 MIC-226 **Al Lith, Inner Fasr, Olimareo**  

**17** - 251    254    MIC-25 **Al Fahal, Yap, normal looking Stylophora**  
**20** - 209 257 258 **Yanbu, Inner Fsar, contains some skinny fingered Stylophora**  
**23** - 176 206 231 243 244 **Yanbu, Al Lith, Al Fahal, almost entirely skinny fingered Stylophora!**  
**31** - 208     227     260     MIC-453 **Yanbu, Al Lith, Innser Fsar, Nakuoro, fatter fingers**  

Looks like we have two types of separation going on here - site and by Stylophora morphology! Pretty cool. 

The Indonesian and Ningaloo samples form nice tight clusters based on site. The Red Sea samples don't cluster by site, but they seem to form clusters based on morphology. Ie. cluster 23 contains almost entirly skinnyer fingered Stylophora.. neat. 

